% ==============================================================================
\section{Discussion and Conclusions}
\label{sec:conc}
% ==============================================================================

% ==============================================================================
\subsubsection*{Discussion of Further Reductions of the Test Effort}
As has been shown in Theorem~\ref{th:maxtracelen}, the maximal length $pq$ of 
traces to be tested for either  failures refinement or trace refinement cannot be further reduced.
It is noteworthy, however, that when testing finite state machines for equivalence, considerably shorter traces can be used. From the classical results published in~\cite{chow:wmethod,vasilevskii1973}, for example, it follows that the maximal trace length to be executed is less or equal to $3p -q$, which is considerably smaller than $pq$ for $p,q\ge 3$.
As a consequence, the investigation of complete test suites establishing failures or trace equivalence is of considerable interest and will be discussed in a future paper.


It is also known from FSM
testing that it is not necessary to check {\it all} traces up to  
length $pq$ 
when testing for reduction of finite state machines (this corresponds to trace refinement). Notable complete strategies supporting this fact have been presented,
for example,
in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}.
From~\cite{Huang2017} it is known that complete FSM testing theories can be
translated to other formalisms, such as Extended Finite State Machines,
Kripke Structures, or CSP, resulting in likewise complete test strategies for
the latter. We intend to study translations of several promising FSM
strategies to CSP in the future. These will effectively reduce the 
upper bound for the number of test executions to be performed, which  has been 
shown to be of order $O\big(\card{\Sigma}^{pq-1}\big)$ for our trace refinement tests 
in Section~\ref{sec:complexity}.  The
bound $\binom{n}{\lfloor n/2\rfloor}$ for the number of sets to be used in probing the SUT for illegal
deadlocks, however, cannot be further reduced, as has been established in
Lemma~\ref{lemma:hseta}.

% ==============================================================================
\subsubsection*{Discussion of Adaptive Test Cases}
The tests suggested
in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were \emph{preset} in the
sense that the trace to be executed was pre-defined for each test. As a consequence,
the authors of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive}
as a third test result, applicable to the situations where the intended trace
of the execution was blocked, due to legal, but nondeterministic behaviour of the
SUT. We consider this as a disadvantage, since, when aiming at executing a specific trace $s$
before being able to check the test objective---for example, the absence of deadlocks when offering a hitting set $H$ of the minimal acceptances---it may take several tries until the full trace $s$ is accepted by the SUT. Considering the complete testing assumption described in Section~\ref{sec:finitecompletefails}, it may even take $c^{\#s}$ tries to reach the end of the trace $s$, if the SUT can legally block every $s$-event due to nondeterminism, so that $c$ 
trials are required for each event is accepted.

%\fixme{From the proofs of the main theorems, and what you say below, I
%thought you did have verdicts coming from deadlock, and you simply chose not
%to mark it with an inc event.}
In  contrast to that, our test cases specified in
Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
U_F(j))$ for failures refinement 
may only stop early with $\epass$ 
after traces $s$ satisfying $\minhits(P/s) = \varnothing$, and may deadlock 
(a) after a trace $s$ where the SUT illegally deadlocks (so $\minhits(P/s) \neq \varnothing$ for
the reference process, but $\minhits(Q/s) = \varnothing$ for the implementation $Q$), or
(b) 
in the final 
step when---just as in the corresponding test cases specified in~\cite{DBLP:conf/icfem/CavalcantiG07}--- hitting sets $H$ are offered to the SUT and it refuses their acceptance.
In both situations (a) and (b) the test fails. As a consequence, far less test repetitions are required according to the complete testing assumption than for the successful execution of all 
test cases specified in~\cite{DBLP:conf/icfem/CavalcantiG07}.

Another distinction of our failures test cases $U_F(j)$ to the tests specified in~\cite{DBLP:conf/icfem/CavalcantiG07} consists in the fact that the former test both trace refinement and
the $conf$ conformance relation in one go, whereas the latter use separate test suites to 
establish these two correctness conditions. Again, we consider the  
structure of the  test cases 
$U_F(j)$ as advantageous, since, when checking acceptance of a hitting set $H$ 
after a trace $s$ for $c$ times according to the complete testing assumption,  
any acceptance of an illegal event $e\in\Sigma - [P/s]^0$ should also be revealed within these $c$ tries.


%
%never block before the final step specified
%by branch (\ref{eq:ufd}), and so we do not need inconclusive test results. It
%should be noted, however, that it is necessary for our test verdicts to
%recognise also deadlocks in the final test step and interpret them as FAIL,
%as described in Section~\ref{sec:finitecompletefails}. In practice, this is
%realised by adding a timeout event to the testing environment which indicates
%deadlock situations. For real-time systems, this is an accepted technique,
%because the SUT has to respond within a pre-defined latency interval,
%otherwise its behaviour is considered to be blocked and regarded as a
%failure. The tests executions $(Q\parallel[\Sigma] U_T(j))$ \fixme{Missing
%hiding.} \fxnote{jp: Here, the hiding operator is not necessary: the possible
%test executions are really the traces of $Q$ parallel $U_T(j)$. The hiding is
%only needed to specify the verdict.}for trace refinement never block at all
%before stopping after the verdict $\epass$ or $\efail$, unless the SUT
%process $Q$ has an unexpected deadlock state. Recall from the explanations
%given in Section~\ref{sec:finitecomplete} that the blocking situation also
%leads to passing the test, because $(\epass\then \Stop)\lessdet_T
%(Q\parallel[\Sigma] U_T(j))\hide\Sigma$ still holds if $(Q\parallel[\Sigma]
%U_T(j))\hide\Sigma$ only produces the empty trace.

%The adaptive behaviour of both our test case types $U_F(j)$ and $U_T(j)$,
%however, induces the obligation to check that {\it all} possible executions
%have been performed before the test can be considered as passed. Typically,
%it is therefore assumed that a \emph{complete testing
%assumption}~\cite{hierons_testing_2004} holds, which means that every
%possible behaviour of the SUT is performed after a finite number of test
%executions. In practice, this is realised by executing each test several
%times, recording the traces that have been performed, and using hardware or
%software coverage analysers to determine whether all possible test execution
%behaviours of the SUT have been observed. Therefore, adaptive test cases come
%at the price of having to apply some grey-box testing techniques enabling us
%to decide whether all SUT behaviours have been observed.

% ==============================================================================
\subsubsection*{Discussion of Fault Domains}
As already mentioned, the work in~\cite{DBLP:conf/pts/CavalcantiS17} defines
a fault domain as the set of processes that refine a given CSP process.  In
that context, only testing for traces refinement is considered, and the
complete test suites may not be finite. So, the work presented here goes well
beyond what is achieved there. On the other hand,
\cite{DBLP:conf/pts/CavalcantiS17} presents an algorithm for test generation
that can be easily adapted to consider additional selection and termination
criteria, like, for example, the length of the traces used to construct
tests. It would be possible, for instance, to use the bound indicated here.
Moreover, specifying a fault domain as a CSP process allows us to model
domain-specific knowledge using CSP. For example, if an initialisation
component defined by a process $I$ can be regarded as correct without further
testing, we can use $I; RUN$ as a fault domain, to indicate that any SUT of
interest implements $I$ correctly, but afterwards has a potentially arbitrary
behaviour specified by $RUN$. In addition, the work
in~\cite{DBLP:conf/pts/CavalcantiS17} is not restricted to finite or
non-terminating processes.

% ==============================================================================
\subsubsection*{Implications for CSP Model Checking}
As explained in the previous sections, passing a test is specified by
$$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$$ for
failures testing. If the SUT $Q$ is not a programmed piece of software or an
integrated hardware or software system, but just another CSP process
specification, it is of course possible to verify the pass criterion using
the FDR4 model checker. For checking the refinement relation $P\lessdet_F Q$,
the pass criterion has to be verified for $j=0,\dots,pq-1$, where $p$ and $q$
indicate the number of nodes in $P$'s transition graph and the maximal number
of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since
the test cases $U_F(j)$ have such a simple structure, it is an interesting
question for further research whether checking $(\epass\then\Stop) \lessdet_F
(Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ for $j=0,\dots,pq-1$ can be faster
than directly checking $P\lessdet_F Q$, as one would do in the usual approach
with FDR4. This is of particular interest, since the checks could be
parallelised on several CPUs. Alternatively it is interesting to investigate
whether the check\footnote{We are grateful to Bill Roscoe for suggesting this
option.}
\[
(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{j=0}^{pq-1} U_F(j)) \hide \Sigma
\]
may perform better than the check of $P\lessdet_F Q$, since the former allows for
other optimisations in the model checker than the latter.

For a large and complex CSP implementation process $Q$, it may be too time consuming to
generate its normalised transition graph, so that its number $q$ of nodes is
unknown. In such a case the suggested options may still be used as efficient
bug finders: use  $p$ of reference process $P$ as initial $q$-value and
increment $q$ from there, as long as each increment reveals new errors.

% ==============================================================================
\subsubsection*{Conclusion}

In this paper, we have introduced finite complete testing strategies for
model-based testing against finite, non-terminating CSP reference models. The
strategies are applicable to the conformance relations failures refinement
and trace refinement. The underlying fault domains have been defined as the
sets of CSP processes whose normalised transition graphs do not have more
than a given number of additional nodes, when compared to the transition
graph of the reference process. For these domains, finite complete test
suites are available. We have shown for the strategy to check failures
refinement that the way of probing the SUT for illegal deadlocks used in our
test cases is optimal in the sense that it is not possible to guarantee
exhaustiveness with fewer probes.

% ==============================================================================

\paragraph{Acknowledgements}
The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for
their advice on using the FDR4 model checker and for very helpful discussions
concerning the potential implications of this paper in the field of model
checking. We are also grateful to Li-Da Tong from National Sun Yat-sen
University, Taiwan, for suggesting the applicability of Sperner's Theorem in
the context of the work presented here. Moreover, we thank Adenilso Simao for
several helpful suggestions. The work of Ana Cavalcanti is funded by the
Royal Academy of Engineering and UK EPSRC Grant EP/R025134/1.
